Detecting skill level and mental templates in Late Acheulean biface morphology: Archaeological and experimental insights

### 2022-07-25 manuscript/manuscript.Rmd
removed the old figures and revised the conclusion.

Regarding the two research questions we proposed in the beginning, our case study suggested that 1) we can delineate the effects of skill level and mental template through the multivariate analysis of morphometric data, where the former is associated with cross-sectional thinning while the latter is reflected in elongation and pointedness; 2) Training has an immediate effect of sharing a common mental template, but 90 hours of training is still not enough for novice to reach the level of expertise as reflected in modern experienced knappers, let alone the Boxgrove foragers. However, it should be noted that it is not our intention to construct a false binary framework and put these two factors as disconnected and opposite concepts. In the future, more robust experimental studies are needed to deepen our understanding of the relationship between skill acquisition and the morphological variability of bifaces as well as their implications for the biological and cultural evolution of the hominin lineages.

>>>>

Regarding the two research questions we proposed in the beginning, our case study suggested that 1) we can delineate the effects of skill level and mental template through the multivariate analysis of morphometric data, where the former is associated with cross-sectional thinning while the latter is reflected in elongation and pointedness; 2) Training has an immediate effect of sharing a common mental template, but 90 hours of training is still not enough for novice to reach the level of expertise as reflected in modern experienced knappers, let alone the Boxgrove tool makers. At a bigger theoretical level it breaks down the distinction between social learning of design targets and individual learning of the skills needed to achieve them. However, it should be noted that it is not our intention to construct a false binary framework and put these two factors as disconnected and opposite concepts. To illustrate, a thin cross section could be part of a mental template or design target and was explicitly instructed by our expert instructor to novices, but novices cannot achieve this technological goal due to the constraint of skill level, making it a robust indicator of the latter. In the future, more robust experimental studies are needed to deepen our understanding of the relationship between skill acquisition and the morphological variability of bifaces as well as their implications for the biological and cultural evolution of the hominin lineages.
